---
title: "Quantitative Trends of Language Access Compliance, 2014-2019"
output: 
  html_notebook:
    toc: true
    toc_float: true
    smooth_scroll: true
---

# Intake Method
```{r Intake Method, echo=TRUE, warning=FALSE, paged.print=TRUE}
method_intake <- X5_intake %>% filter(Department != "Test Department") %>% group_by(Year) %>% filter(Intake == TRUE) %>% count(Intake)
method_survey <-X5_intake %>% filter(Department != "Test Department") %>% group_by(Year) %>% filter(Survey == TRUE) %>% count(Survey)
method_tele <- X5_intake %>% filter(Department != "Test Department") %>% group_by(Year) %>% filter(Telephonic == TRUE) %>% count(Telephonic)

temp <- merge(method_intake, method_survey, by = "Year") %>% select(Year, n.x, n.y) %>% rename(Intake = n.x, Survey = n.y)
method <- merge(temp, method_tele, by = "Year") %>% select(Year, Intake, Survey, n) %>% rename(Telephonic = n) 
method_pivot <- pivot_longer(method, !Year, names_to = "Method", values_to = "count")
```

```{r Intake Method — Graph, echo=TRUE}
ggplot(data = method_pivot, aes(x=Year, y = count, group = Method)) + 
  geom_line(aes(color = Method, size = Method, linetype = Method)) +
  geom_point() + 
  scale_size_manual(values = c(1.5, 0.8, 0.8)) + 
  scale_linetype_manual(values=c("solid", "twodash", "twodash")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Method for Counting Client Interactions") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```

# Client Information {.tabset}

This information summarizes already-aggregated information from departments.
```{r Reported LEP Total, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
client_total <- X6_client %>% group_by(Year) %>% summarize(total = sum(LEP.Total, na.rm = TRUE)) %>% filter(!is.na(Year))
```

## All Interactions (Reported Total)
This information summarizes all of the reported LEP numbers per language, in order to comeup with our own summary. We use this to check the validity of the total LEP client numbers departments provide to us.

```{r Graph - Reported LEP Client Total, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = client_total, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Client Interactions") +
  ylab("Number of Total Client Interactions") + 
  xlab("Fiscal Year")
```

## All Interactions (Manual Total)
```{r Graph - Autosummed LEP Client Total, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
client_total_manual <- X6_client %>% group_by(Year) %>% summarize(total = sum(LEP.CAN + LEP.MAN + LEP.SPA + LEP.FIL + LEP.VIET + LEP.RUS + LEP.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = client_total_manual, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Client Interactions") +
  ylab("Number of Total Client Interactions") + 
  xlab("Fiscal Year")
```

## By Language
```{r Graph - LEP Clients by Language, echo=TRUE}
clients_language <- X6_client %>% group_by(Year) %>% summarize(CAN = sum(LEP.CAN, na.rm = TRUE), MAN = sum(LEP.MAN, na.rm = TRUE), 
                                           SPA = sum(LEP.SPA, na.rm = TRUE), FIL = sum(LEP.FIL, na.rm = TRUE),
                                           RUS = sum(LEP.RUS, na.rm = TRUE), VIET = sum(LEP.VIET, na.rm = TRUE),
                                           OTH = sum(LEP.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))
clients_pivot <- pivot_longer(clients_language, !Year, names_to = "Language", values_to = "count")

ggplot(data = clients_pivot, aes(x=Year, y = count, group = Language)) + 
  geom_line(aes(color = Language)) +
  geom_point(aes(color = Language)) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Client Interactions over Time by Language") +
  ylab("Number of Client Interactions") + 
  xlab("Fiscal Year")
```
# Translations {.tabset}

## All Translations

This information looks at the total number of translations over the years.
```{r Translations Provided, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
translations <- X7_translations %>% group_by(Year) %>%
  summarize(total = sum(Translations.CHI + Translations.SPA +
                          Translations.FIL + Translations.VIET +
                          Translations.RUS + Translations.OTHER,
                        na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = translations, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Translations") +
  ylab("Number of Translations") + 
  xlab("Fiscal Year")
```

## By Language
```{r Translations - Language, echo=TRUE}
translations_language <- X7_translations %>% group_by(Year) %>% summarize(CAN = sum(Translations.CHI, na.rm = TRUE), 
                                           SPA = sum(Translations.SPA, na.rm = TRUE), FIL = sum(Translations.FIL, na.rm = TRUE),
                                           RUS = sum(Translations.RUS, na.rm = TRUE), VIET = sum(Translations.VIET, na.rm = TRUE),
                                           OTH = sum(Translations.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))
translations_pivot <- pivot_longer(translations_language, !Year, names_to = "Language", values_to = "count")

ggplot(data = translations_pivot, aes(x=Year, y = count, group = Language)) + 
  geom_line(aes(color = Language)) +
  geom_point(aes(color = Language)) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Translations over Time by Language") +
  ylab("Number of Translations by Language") + 
  xlab("Fiscal Year")
```
# Interpretations - Phone {.tabset}

## All Interpretations

```{r Phone Interpretations, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
interpret_phone <- X8_interpret_phone %>% group_by(Year) %>%
  summarize(total = sum(Calls.CAN + Calls.MAN + Calls.SPA +
                          Calls.FIL + Calls.VIET + Calls.RUS +
                          Calls.OTHER,
                        na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = interpret_phone, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Total Phone Interpretations over Time") +
  ylab("Number of Phone Interpretations") + 
  xlab("Fiscal Year")
```

## By Language
```{r Phone Interpretations - Language, echo=TRUE}
interpret_phone_language <- X8_interpret_phone %>% group_by(Year) %>% summarize(CAN = sum(Calls.CAN, na.rm = TRUE), MAN = sum(Calls.MAN, na.rm = TRUE), 
                                           SPA = sum(Calls.SPA, na.rm = TRUE), FIL = sum(Calls.FIL, na.rm = TRUE),
                                           RUS = sum(Calls.RUS, na.rm = TRUE), VIET = sum(Calls.VIET, na.rm = TRUE),
                                           OTH = sum(Calls.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

interpret_phone_pivot <- pivot_longer(interpret_phone_language, !Year, names_to = "Language", values_to = "count")

ggplot(data = interpret_phone_pivot, aes(x=Year, y = count, group = Language)) + 
  geom_line(aes(color = Language)) +
  geom_point(aes(color = Language)) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Phone Interpretations over Time by Language") +
  ylab("Number of Phone Interpretations") + 
  xlab("Fiscal Year")
```

# Interpretations - In-Person {.tabset}

## All Interpretations
```{r Total - Graph, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
interpret_person <- X9_interpret_person %>%
  group_by(Year) %>%
  summarize(total = sum(Interpretations.CAN + Interpretations.MAN +
                          Interpretations.SPA + Interpretations.FIL +
                          Interpretations.VIET + Interpretations.RUS +
                          Interpretations.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = interpret_person, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  ylim(0, 150000) +
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Total In-Person Interpretations over Time") +
  ylab("Number of In-Person Interpretations") + 
  xlab("Fiscal Year")
```

## By Language
```{r In-Person Language Graph, echo=TRUE}
interpret_person_language <- X9_interpret_person %>% group_by(Year) %>% summarize(CAN = sum(Interpretations.CAN, na.rm = TRUE), 
                                                                                  MAN = sum(Interpretations.MAN, na.rm = TRUE), 
                                           SPA = sum(Interpretations.SPA, na.rm = TRUE), FIL = sum(Interpretations.FIL, na.rm = TRUE),
                                           RUS = sum(Interpretations.RUS, na.rm = TRUE), VIET = sum(Interpretations.VIET, na.rm = TRUE),
                                           OTH = sum(Interpretations.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

interpret_person_pivot <- pivot_longer(interpret_person_language, !Year, names_to = "Language", values_to = "count")

ggplot(data = interpret_person_pivot, aes(x=Year, y = count, group = Language)) + 
  geom_line(aes(color = Language)) +
  geom_point(aes(color = Language)) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "In-Person Interpretations over Time by Language") +
  ylab("Number of In-Person Interpretations") + 
  xlab("Fiscal Year")
```

# Staffing - All Multilingual Staff {.tabset}

## All Staff
```{r Multilingaul Staff - All Graph, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
staffing <- X11_staffing %>% group_by(Year) %>% summarize(total = sum(Staff.Total.Multilingual, na.rm = TRUE)) %>% filter(!is.na(Year))
staff_capacity <- X11_staffing %>% group_by(Year) %>% summarize(total = sum(Staff.CAN + Staff.MAN + Staff.SPA + Staff.FIL + Staff.VIET + Staff.RUS +
                          Staff.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = staffing, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  ylim(0, 6500) +
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Total Multilingual Staff over Time") +
  ylab("Number of Staff") + 
  xlab("Fiscal Year")
```

## By Language
```{r Multilingual Staff - Language Graph, echo=TRUE}
staffing_language <- X11_staffing %>% group_by(Year) %>% summarize(CAN = sum(Staff.CAN, na.rm = TRUE), MAN = sum(Staff.MAN, na.rm = TRUE), 
                                           SPA = sum(Staff.SPA, na.rm = TRUE), FIL = sum(Staff.FIL, na.rm = TRUE),
                                           RUS = sum(Staff.RUS, na.rm = TRUE), VIET = sum(Staff.VIET, na.rm = TRUE),
                                           OTH = sum(Staff.OTHER, na.rm = TRUE)) %>% filter(!is.na(Year))

staffing_pivot <- pivot_longer(staffing_language, !Year, names_to = "Language", values_to = "count")

ggplot(data = staffing_pivot, aes(x=Year, y = count, group = Language)) + 
  geom_line(aes(color = Language)) +
  geom_point(aes(color = Language)) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Multilingual Staff over Time by Language") +
  ylab("Number of Multilingual Staff") + 
  xlab("Fiscal Year")
```

# Staffing - Certified Multilingual Staff
```{r Certified Staff Graph, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
staff_cert <- X11_staffing %>% group_by(Year) %>% summarize(total = sum(Staff.Total.Multilingual.Cert, na.rm = TRUE)) %>% filter(!is.na(Year))

ggplot(data = staff_cert, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  ylim(0, 3000)
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Total Certified Multilingual Staff over Time") +
  ylab("Number of Staff") + 
  xlab("Fiscal Year")
```

# Department Certifications for Multilingual Staff

```{r Certification Data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
cert <- X12_bilingual_cert %>% group_by(Year) %>%
  count(Cert.DHR, Cert.External, Cert.Other, Cert.No.Method)

cert.dhr <- cert %>% filter(Cert.DHR == TRUE) %>% group_by(Year) %>% summarise(total = sum(n))
cert.ext <- cert %>% filter(Cert.External == TRUE) %>% group_by(Year) %>% summarise(total = sum(n))
cert.oth <- cert %>% filter(Cert.Other == TRUE) %>% group_by(Year) %>% summarise(total = sum(n))

temp <- merge(cert.dhr, cert.ext, by = "Year") %>% select(Year, total.x, total.y) %>% rename(DHR = total.x, External = total.y)
cert_all <- merge(temp, cert.oth, by = "Year") %>% select(Year, DHR, External, total) %>% rename(Other = total) 
cert_pivot <- pivot_longer(cert_all, !Year, names_to = "Method", values_to = "count")

```

```{r Certification Graph, echo=TRUE}
ggplot(data = cert_pivot, aes(x=Year, y = count, group = Method)) + 
  geom_line(aes(color = Method, linetype = Method, size = Method)) +
  geom_point() + 
  scale_size_manual(values = c(0.8, 0.8, 0.8)) + 
  scale_linetype_manual(values = c("solid", "solid", "solid")) +
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Multilingual Certification Methods") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```
# Department Language Policies
```{r LAO Policy Graph, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
lao_yes <- X13_dep_lao %>% group_by(Year) %>% filter(Department.Has.Language.Policy == "Yes") %>% count(Department.Has.Language.Policy)

ggplot(data = lao_yes, aes(x=Year, y = n, group = 1)) + 
  geom_line() +
  geom_point() + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Departments with Internal Language Policies") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```

# Department Asessment of Language Services {.tabset}

## Employees
```{r Staffing, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
staff_yes <- X14_services_assess %>% group_by(Year) %>% filter(Curr.Bilingual.Employees.Adequate == "Yes") %>%  count(Curr.Bilingual.Employees.Adequate)

ggplot(data = staff_yes, aes(x=Year, y = n, group = 1)) + 
  geom_line() +
  geom_point() + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Departments with Adequate Number of Multilingual Employees") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```

## Recorded Messages
```{r Messages, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
messages_yes <- X14_services_assess %>% group_by(Year) %>% filter(!is.na(Availability.Recorded.Messages)) %>% count(Availability.Recorded.Messages)
ggplot(data = messages_yes, aes(x=Year, y = n, group = Availability.Recorded.Messages)) + 
  geom_line(aes(color = Availability.Recorded.Messages, linetype = Availability.Recorded.Messages, size = Availability.Recorded.Messages)) +
  geom_point() + 
  scale_size_manual(values = c(0.8, 1.5, 0.8)) + 
  scale_linetype_manual(values=c("twodash", "solid", "twodash")) + 
  scale_color_manual(values=c('#32CD32', '#008000', '#FF4500')) +
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Quality of Available Recorded Messages") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```

## LEP Notices
```{r LEP Notices, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
notices <- X14_services_assess %>% group_by(Year) %>% filter(!is.na(Quality.LEP.Notices)) %>% count(Quality.LEP.Notices)

ggplot(data = notices, aes(x=Year, y = n, group = Quality.LEP.Notices)) + 
  geom_line(aes(color = Quality.LEP.Notices, linetype = Quality.LEP.Notices, size = Quality.LEP.Notices)) +
  geom_point() + 
  scale_size_manual(values = c(0.8, 1.5, 0.8)) + 
  scale_linetype_manual(values=c("twodash", "solid", "twodash")) + 
  scale_color_manual(values=c('#32CD32', '#008000', '#FF4500')) +
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Quality of LEP Notices") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```

## Translation Services
```{r Translation Services, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
translation_services <- X14_services_assess %>% group_by(Year) %>% filter(!is.na(Translation.Services.Procedure)) %>% count(Translation.Services.Procedure)

ggplot(data = translation_services, aes(x=Year, y = n, group = Translation.Services.Procedure)) + 
  geom_line(aes(color = Translation.Services.Procedure, linetype = Translation.Services.Procedure, size = Translation.Services.Procedure)) +
  geom_point() + 
  scale_size_manual(values = c(0.8, 1.5, 0.8)) + 
  scale_linetype_manual(values=c("twodash", "solid", "twodash")) + 
  scale_color_manual(values=c('#32CD32', '#008000', '#FF4500')) +
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Quality of Internal Translation Services Procedure") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")

```

## Interpretation Services
```{r Interpretation Procedures, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
interpret_services <- X14_services_assess %>% group_by(Year) %>% filter(!is.na(Interpretation.Procedures)) %>% count(Interpretation.Procedures)

ggplot(data = interpret_services, aes(x=Year, y = n, group = Interpretation.Procedures)) + 
  geom_line(aes(color = Interpretation.Procedures, linetype = Interpretation.Procedures, size = Interpretation.Procedures)) +
  geom_point() + 
  scale_size_manual(values = c(0.8, 1.5, 0.8)) + 
  scale_linetype_manual(values=c("twodash", "solid", "twodash")) + 
  scale_color_manual(values=c('#32CD32', '#008000', '#FF4500')) +
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Quality of Internal Interpretation Services Procedure") +
  ylab("Number of Departments") + 
  xlab("Fiscal Year")
```
# Department Language Services Budget
```{r Budget Analysis & Graph, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
budget <- X18_budget %>%
  group_by(Year) %>%
  summarize(total = sum(Lang.Services.Budget, na.rm = TRUE))

budget <- budget[-c(7),]

ggplot(data = budget, aes(x=Year, y = total, group = 1)) + 
  geom_line(na.rm = TRUE) +
  geom_point() + 
  scale_size_manual(values = c(1.5)) + 
  scale_linetype_manual(values=c("solid")) + 
  scale_x_discrete(labels = c("2014-16", "2015-16", "2016-17", "2017-18", "2018-19", "2019-20")) +
  labs(title = "Total Language Service Budget over Time") +
  ylab("Budget, $") + 
  xlab("Fiscal Year")

```